Practical Optional Types for Clojure

### 2015-02-28 formal-model.tex
meta

\paragraph{Class path element} Our second path element \classpe{} is used in the latent
object of the constant \classconst{} function. Like Clojure's \clj{class}
function \classconst{} returns the argument's class or \nil{}
if passed \nil{}.
$$
\begin{array}{lrlr}
  \pesyntax{}   &::=& \ldots \alt {\classpe{}}
                &\mbox{Path Elements}
\end{array}
$$
\begin{mathpar}
\constanttypefigure{}
\end{mathpar}
The full semantics are given in \figref{main:figure:primitivesem}.
The definition of \updateliteral{} supports various idioms relating to \classpe{}
which we introduce in \secref{sec:isaformal}.

>>>>

The full definition of \updateliteral{} is given in \figref{main:figure:update}
which considers both keys a path elements as well as the \classconst{}
path element described below.
In the absence of paths, update simply performs set-theoretic operations
on types; see \figref{main:figure:restrictremove} for details.

\paragraph{Class path element} Our second path element \classpe{} is used in the latent
object of the constant \classconst{} function. Like Clojure's \clj{class}
function \classconst{} returns the argument's class or \nil{}
if passed \nil{}.
$$
\begin{array}{lrlr}
  \pesyntax{}   &::=& \ldots \alt {\classpe{}}
                &\mbox{Path Elements}
\end{array}
$$
\begin{mathpar}
\constanttypefigure{}
\end{mathpar}
The dynamic semantics are given in \figref{main:figure:primitivesem}.
The definition of \updateliteral{} supports various idioms relating to \classpe{}
which we introduce in \secref{sec:isaformal}.

### 2015-02-28 metatheory.tex
formatting
meta

Our proof for type soundness is similar to~\citet{TF10}. We add
errors and \wrong{} value and prove
well-typed programs do not go wrong.

We make our assumptions about Java explicit. We concede that
method and constructor calls may diverge or error, but we assume they can
never go wrong.

@@

We use an extra lemma to support our main soundness lemma. Consistency
ensures that occurrence typing does not refer to variables
hidden inside a closure.

@@

{\wrongtheoremnoproof{main}}

Null-pointer exceptions are ruled-out in typed code.

{\nilinvoketheoremnoproof{main}}

>>>>

We prove type soundness follow using the same technique as~\citet{TF10}. 
We also include errors and a \wrong{} value and prove
well-typed programs do not go wrong.

Rather than modelling Java's dynamic semantics, we instead
make our assumptions about Java explicit. We concede that
method and constructor calls may diverge or error, but we assume they can
never go wrong. (Assumptions for other operations are given in the supplemental
material).

@@

For the purposes of our soundness proof, we require that all values
are \emph{consistent}.
Consistency
ensures that occurrence typing does not refer to variables
hidden inside a closure.

@@

{\wrongtheoremnoproof{main}}
%
%{\nilinvoketheoremnoproof{main}}
